
---
title: "Reproducible Research: Peer Assessment 1"
output: 
html_document:
keep_md: true
---
Loading the knitr package
```{r}
library(knitr)
```

Setting Global Option where echo = true so that someone will be able to read the code.
```{r setoptions, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE, results = "asis")
```


## Loading and preprocessing the data
```{r, results='markup', warning=TRUE, message=TRUE}
if (!file.exists("activity.csv") )
{
  url <- 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'  
  download.file(url, destfile='repdata-data-activity.zip', mode='wb')  
  unzip('repdata-data-activity.zip')
}

data <- read.csv("activity.csv", header = TRUE, sep = ",", na.strings = "NA", colClasses=c("numeric", "character", "numeric"))

```


## What is mean total number of steps taken per day?

1. Make a histogram of the total number of steps taken each day


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
