
---
title: "Reproducible Research: Peer Assessment 1"
output: 
html_document:
keep_md: true
---
Loading the knitr package
```{r}
library(knitr)
```

Setting Global Option where echo = true so that someone will be able to read the code and results.
```{r setoptions, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE, results = "asis")
```


## Loading and preprocessing the data
```{r, LoadingAndProcessingData}
if (!file.exists("activity.csv") )
{
  url <- 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'  
  download.file(url, destfile='repdata-data-activity.zip', mode='wb')  
  unzip('repdata-data-activity.zip')
}

data <- read.csv("activity.csv", header = TRUE, sep = ",", na.strings = "NA", colClasses=c("numeric", "character", "numeric"))

```


## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

```{r, StepsByDayUsingAggregate}
steps_by_day <- aggregate(steps ~ date, data=data, sum)
```

2. Make a histogram of the total number of steps taken each day
```{r, HistogramStepsPerDay}
hist(steps_by_day$steps, main = paste("Total Steps Per Day"), col="red",xlab="Number of Steps")
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r, MeanMedianTotalSteps}
meanSteps <- mean(steps_by_day$steps)
meanSteps

medianSteps <- median(steps_by_day$steps)
medianSteps
```

###### The mean total number of steps taken per day is `r meanSteps` and the median total number of steps taken per day `r medianSteps`.



## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
