
---
title: "Reproducible Research: Peer Assessment 1"
output: PA1_template.Rmd
html_document:PA1_template.html
keep_md: true
---
Loading the knitr package
```{r}
library(knitr)
```

Setting Global Option where echo = true so that someone will be able to read the code and results.
```{r setoptions, echo=TRUE}
knitr::opts_chunk$set(echo = TRUE, results = "asis")
```


## Loading and preprocessing the data
```{r, LoadingAndProcessingData}
if (!file.exists("activity.csv") )
{
  url <- 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip'  
  download.file(url, destfile='repdata-data-activity.zip', mode='wb')  
  unzip('repdata-data-activity.zip')
}

data <- read.csv("activity.csv", header = TRUE, sep = ",", na.strings = "NA", colClasses=c("numeric", "character", "numeric"))

```


## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

```{r, StepsByDayUsingAggregate}
steps_by_day <- aggregate(steps ~ date, data=data, sum)
```

2. Make a histogram of the total number of steps taken each day
```{r, HistogramStepsPerDay}
hist(steps_by_day$steps, breaks = 20, main = paste("Total Steps Per Day"), col="red",xlab="Number of Steps")
```



3. Calculate and report the mean and median of the total number of steps taken per day
```{r, MeanMedianTotalSteps}
meanSteps <- mean(steps_by_day$steps)
meanSteps

medianSteps <- median(steps_by_day$steps)
medianSteps
```

###### The mean total number of steps taken per day is `r meanSteps` and the median total number of steps taken per day `r medianSteps`.



## What is the average daily activity pattern?

1. Make a time series plot (i.e. type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

Load lattice package and prepare data for plot
```{r, TimeSeriesPlot}

library(lattice)

intervalSteps <- aggregate(x= list(steps = data$steps), by =list(interval = data$interval), FUN = mean, na.rm = TRUE)

xyplot(intervalSteps$steps ~ intervalSteps$interval, type = "l", xlab ="5-minute interval", ylab = "the average number of steps taken", main = "Average daily activity pattern")

```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r, maxIntervalSteps}
maxIntervalSteps <- intervalSteps[which.max(intervalSteps$steps),]
maxIntervalSteps
```

###### The 5 minute interval with maximum number of steps is at `r maxIntervalSteps$interval` interval with the maximum of `r maxIntervalSteps$steps` steps.


## Imputing missing values
1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r, missingSteps}
missingSteps <- sum(is.na(data))
```
###### The total number of missing values in the dataset is `r missingSteps`.


2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
```{r, newStepsStrategy}
newData <- function(data){
        for (i in 1:nrow(data)){
                if(is.na(data[i,1])){
                        data[i,"steps"] <- intervalSteps[intervalSteps$interval == data[i,"interval"],"steps"] 
                }
        }
        return(data)
}
```


3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r, NewStepsData }
filledData <- newData(data)
```

4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r, NewDataHistogram}
newSteps <- aggregate(steps ~ date, data=filledData, sum)

hist(newSteps$steps, breaks = 20, main = paste("Total Steps Each Day (New)"), col="red",xlab="Number of Steps")
```

```{r, NewMeanMedianSteps}
newMeanSteps <- mean(newSteps$steps)
newMeanSteps

newMedianSteps <- median(newSteps$steps)
newMedianSteps

diffMedianSteps <- newMedianSteps - medianSteps 
diffMedianSteps
```

The new mean steps is `r newMeanSteps` and the new median steps is `r newMedianSteps`

Judging from the new values, there are no difference in mean steps but a difference in median steps by `r diffMedianSteps`


## Are there differences in activity patterns between weekdays and weekends?
1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r, filledDataWithDays}
library(lubridate)
filledData$day <- weekdays(as.Date(filledData$date))
filledData$daytype <- ifelse(weekdays(as.Date(filledData$date)) %in% c("Saturday","Sunday"), "weekend", "weekday")
str(filledData)
table(filledData$daytype)
```

2. Make a panel plot containing a time series plot (i.e. type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r, WeekdayWeekendPlot}
meanSteps <- aggregate(filledData$steps, list(as.numeric(filledData$interval), filledData$daytype), FUN = "mean")

names(meanSteps) <- c("interval","weekDays", "avgSteps")

xyplot(meanSteps$avgSteps ~ meanSteps$interval | meanSteps$weekDays, 
       layout = c(1, 2), type = "l", 
       xlab = "Interval", ylab = "Number of steps", main = "5 mins interval on average number of steps taken")
```
